---
title: "ColombiAPI: Exploring Colombia's Public Data with API-Colombia"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{ColombiAPI: Exploring Colombia's Public Data with API-Colombia}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(ColombiAPI)
library(dplyr)

```


# Introduction 

The `ColombiAPI` package provides a seamless interface to access diverse public data about `Colombia` through the `API-Colombia`, a RESTful API.
The package enables users to explore various aspects of `Colombia`, including general information, geography, and cultural insights.
It includes five API-related functions to retrieve data on topics such as `Colombia`'s general information, airports, departments, regions, and presidents.

Additionally, `ColombiAPI` offers a built-in function to view the datasets available within the package.
The package also includes curated datasets covering Bogota air stations, business and holiday dates, public schools, Colombian coffee exports,
cannabis licenses, Medellin rainfall, and malls in Bogota, making it a comprehensive tool for exploring `Colombia`'s data.

## Functions for API-Colombia

The `ColombiAPI` package provides five primary functions to retrieve data from the `API-Colombia`,
we'll show you how two of them work in this section:


- `get_Colombia_info()`
Retrieves general information about `Colombia`, including its capital, surface area, population, and languages.

```{r colombia_info}

get_Colombia_info()


```


-  `get_regions_list()`
Retrieves information about `Colombia`'s regions, showcasing their unique characteristics.

```{r colombia_regions}

get_regions_list()


```



- `get_departments_list()`: Lists all the departments of `Colombia`, including their geographical information.


- `get_presidents_list()`:  Provides a list of all presidents in `Colombia`'s history.


 - `get_airports_list()`:  Get the List of Airports in `Colombia`.



## Datasets Included in ColombiAPI

In addition to API functions, `ColombiAPI` includes several preloaded datasets that provide valuable insights into various aspects of `Colombia`:

- `Bogota_airstations_df`: A data frame containing coordinates of air quality monitoring stations in Bogota.

- `Bogota_business_Date`: A Date object listing official business dates in Bogota. 

- `Bogota_holidays_Date`: A Date object listing official holidays in Bogota. 

- `Bogota_malls_tbl_df`: A tibble with detailed information about malls in Bogota. 

- `Cannabis_Licenses_tbl_df`: A tibble containing data on cannabis licenses issued in Colombia. 

- `Colombia_coffee_tbl_df`: A tibble with details of Colombian coffee export and import for the year 2016.

- `Medellin_rain_tbl_df`: A tibble providing rainfall data for Medellin. 

- `Tulua_Public_Schools_tbl_df`: A tibble with information on public schools in Tulua. 

### Dataset Suffixes

Each dataset in `ColombiAPI` is labeled with a `suffix` to indicate its type and structure:

- `_df`: A standard data frame.

- `_tbl_df`: A tibble, a modern version of a data frame with better formatting and functionality.

- `_Date`: An object of type Date.

## Viewing All Available Datasets 

The `ColombiAPI` package includes a built-in function, `view_datasets()`, which allows users to view a comprehensive list of all datasets available in the package directly from the console:

```{r view_datasets}

view_datasets()

```


## Conclusion

The `ColombiAPI` package offers a powerful set of tools for accessing and analyzing public data about Colombia.
From exploring general information about the country to diving into specific datasets on topics like air quality, public schools, or coffee exports, `ColombiAPI` enables users to engage with Colombia's rich data landscape easily.
For more detailed examples and use cases, refer to the full package documentation or explore the datasets using the built-in functions.

<div class="tocify-extend-page" data-unique="tocify-extend-page" style="height: 0;"></div>
